---
title: "R Notebook"
output: html_notebook
---
--- Reading in the Dataset and the correct package for desired tools
```{r}
library(tidyverse)
books <- read_csv("data/books.csv")
```

--- Investigating the Data
```{r}
glimpse(books)
str(books)
view(books)
dim(books)
# 11123 arguments and 13 observations, made up of <chr> and <dbl>
# have noticed that some books may have been removed from the dataset, as                                                             "rowid" and "bookID" do not match in terms of similar numbers
# also some arguments have inconsistent layout

books %>% 
  summarise(across(everything(), 
                   .fns = ~ sum(is.na(.x))))
# details if any NA within data and where they are - when run shows no NA's
```

--- Cleaning the data using Janitor Package 
```{r}
library("janitor")

books <- books %>% 
  clean_names()
# removed capitalisation of titles and put in _'s
```

--- Q1. The top 5 rated authors?
```{r}
best_authors <- books %>% 
  group_by(authors) %>% 
  summarise(avg_rating = mean(average_rating))

slice_max(best_authors, avg_rating, n = 5, with_ties = FALSE)
# after using this, I have noticed that this is an unfair comparison as some                                                          books have been rated very little and received higher ratings
```

```{r}
avg_rating_count <- books %>% 
  summarise(avg_rating_count = mean(ratings_count)) %>% 
  pull()
avg_rating_count
# aiming to get the average number of ratings per book, to use that figure as                                                             a limit to the number of books I will make comparisons with
```

```{r}
books %>% 
  filter(ratings_count >= avg_rating_count) %>% 
  group_by(authors) %>% 
  summarise(avg_rating = mean(average_rating)) %>% 
  slice_max(avg_rating, n = 5)
# had to group the authors, then get an average of their ratings, then slice
# Bill Watterson according to this data set it the highest rated author
```

```{r}
books %>% 
  filter(authors == "Bill Watterson")
# finding out what books the top rated author wrote - Calvin And Hobbes seems                                                           to be the predominant book, which is a popular American strip
```


--- Q2. The top 5 rated books?

```{r}
books %>% 
  filter(ratings_count >= avg_rating_count ) %>% 
  slice_max(average_rating, n = 5 )
# again, had to use ratings higher than the avg rating for fairness
# 3 Calvin And Hobbes with 2 Harry Potter books
```

--- Q3. Longest book and shortest book?
```{r}
books %>%
  slice_max(num_pages, n = 5) %>% 
  summarise(title,num_pages)
# a book of books is the longest, therefore changed to show top 5
# The Second World War shows second longest, as a single book this is a fairer                                                            answer

books %>%
  slice_min(num_pages, n = 5) %>% 
  summarise(title,num_pages)
# after investigating more of the data, it appears inaccurate enough that you                                                             cannot deduce the shortest book without using external resources
```




--- Q4. How many books are written in each language code?
```{r}
books %>% 
  group_by(language_code) %>% 
  summarise(count = n())
# this showed that there are multiple types of english, so wanted to group them                                                           together under one title
```

Used this code from a previous example of editing the data to help with the below code.

  students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ "Other number"
    )
  )

```{r}
books %>% 
    mutate(
    language_code = case_when(
    language_code == "en-CA" ~ "eng",
    language_code == "en-GB" ~ "eng",
    language_code == "en-US" ~ "eng",
    language_code == "enm"   ~ "eng",
      TRUE                   ~ language_code
    )
  ) %>% 
  group_by(language_code) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

--- Q5. The publisher with the highest average ratings?
```{r}
books %>% 
  filter(ratings_count >= avg_rating_count) %>% 
  group_by(publisher) %>% 
  summarise(avg_rating = mean(average_rating)) %>% 
  slice_max(avg_rating)
# Andrews McMeel Publishing is the publisher of Calvin And Hobbes - written by                                                            the highest rated author!
```


--- (Neringa's q) Q6. Find 10 publishers with the highest amount of books, and for each of them
find their highest rated book?
```{r}
highest_number_books <- books %>% 
  group_by(publisher) %>% 
  summarise(number_of_books = n()) %>% 
  slice_max(number_of_books, n = 10)

highest_number_books %>% 
  left_join(books, by = "publisher") %>% 
  group_by(publisher) %>% 
  slice_max(average_rating, n = 1)
```




--- (Howard's q) 
```{r}
books %>% 
        group_by(publisher) %>%               
        slice_max(ratings_count, n = 1, with_ties = FALSE) %>%
        select (publisher,ratings_count) %>%
        arrange(desc(ratings_count))  %>%
        head(10)
```

