---
title: "Programming in R Homework - Answers"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

# MVP

<br> 

---**Question 1.** \
Find the weather forecast for the next three days online. Create a list of lists in which each inner list should be labelled with the name of the day of the week. Each inner list should contain named elements telling you the temperature, chance of rain and wind speed.

```{r}
weather_forecast <- list(
  "Thursday" = list(temp_celicious = 7, chance_of_rain_percent = 91, wind_speed_mph = 11),
  "Friday"   = list(temp_celicious = 6, chance_of_rain_percent = 10, wind_speed_mph = 15),
  "Saturday" = list(temp_celicious = 8, chance_of_rain_percent = 80, wind_speed_mph = 11)
)
weather_forecast
``

<br>

---**Question 2.** \
Using the list you've just made find:

  i. The chance of rain tomorrow
```{r}
weather_forecast$Thursday$chance_of_rain_percent
```
  
  ii. The wind speed on the day after tomorrow
```{r}
weather_forecast$Friday$wind_speed_mph
```
  
  iii. The temperature tomorrow.
```{r}
weather_forecast$Thursday$temp_celicious
```
  

<br>

---**Question 3.** \
Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days.

```{r}
fortnight_weather_forecast <-
  matrix(c(7,6,8,14,12,10,10,10,9,8,8,7,7,7), nrow = 2, ncol= 7, byrow = TRUE)

fortnight_weather_forecast
```


<br>

---**Question 4.** \
From the matrix you just made select:

  i. The third day of the second week
```{r}
fortnight_weather_forecast[2,3]
```
  
  ii. The whole second week
```{r}
fortnight_weather_forecast[2,1:7]
```
  
  iii. The first, third and fifth days in both weeks.

```{r}
fortnight_weather_forecast[,c(1,3,5)]
```

<br>

---**Question 5.** \
Write a function that pre-appends "The answer is: " to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return "Invalid input".

```{r}

# my answer
task_function_5 <- function(x){
  if (is.character(x)) {
    if x <10{
      if x > 10{
        print("The answer is: " x)  
       }
       print(x)
    }
    print(x / 10)
  }
}  

# instructor answer
answer_or_divide <- function(x) {
  if (is.character(x)) {
    return(paste("The answer is: ", x))
  }

  if (is.numeric(x)) {
    if (x < 10) {
      return(x)
    } else {
      return(x / 10)
    }
  }

  return("Invalid input")
}
```

<br>

---**Question 6.** \
Write a function which checks if two vectors have the same elements. 

* If they have the same elements in the same order, return "Exactly the same"
* If they have the same elements but in a different order, return "The same"
* If they don't have the same elements, return "Different".

```{r}
# my answer

task_function_6 <- function(x = c(), y = c()){ # no need to specifcy that x + y are vecotrs
  if (x == y){ # have to use all() function to get a single TRUE\FALSE when the == checks if each vector is the same 
    if (sort(x == y)){ # when using sort() need to sort each variable individually
      print("Exactly the same")
    }
  print("The same")
  } else {
  print("Different")
  }
}

# instructor answer
is_same_vector <- function(vector_1, vector_2) {
  if (all(vector_1 == vector_2)) {
    return("Exactly the same")
  }

  if (all(sort(vector_1) == sort(vector_2))) {
    return("The same")
  }

  return("Different")
}
```
`

<details>
<summary>**Hints**</summary>
* You can check if two vectors have the same elements but in a different order by using `sort` to sort each vector first and then checking if they are equal.
* Have a look at the `all()` function
</details>
</blockquote>


# Extension

<br>

---**Question 7.** \
Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

```{r}
num_rows <- dim(product_matrix)[1]
num_cols <- dim(product_matrix)[2]

for(row_index in 1:num_rows){
  for(col_index in 1:num_cols){
    product_matrix[row_index,col_index] <- row_index * col_index
  }
}

product_matrix

# create empty matrix
mult_table <- matrix(nrow = 12, ncol = 12)

for (outer_factor in 1:12){
  for (inner_factor in 1:12){
    mult_table[outer_factor, inner_factor] = outer_factor * inner_factor
  }
}

mult_table
```


*From: https://adriann.github.io/programming_problems.html*

<br>

---**Question 8.** \
Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function `cat()`. This is similar to `print()`, but prints directly, without vector markers. You can also include newlines `"\n"` when using `cat()`.

```
1 | ***
2 | *
3 | *******
4 | ****
5 | **
```

```{r}
make_histogram <- function(x){
  
  distinct <- sort(unique(x))
  
  for (i in distinct){
    
    count <- sum(x == i)
    
    cat(i)
    cat(" | ")
    cat(rep("*", count))
    cat("\n")
    
  }
  
}
```

```{r}
numbers <- sample(1:5, size = 50, replace = TRUE) 
numbers

make_histogram(numbers)
```


