---
title: "R Notebook"
output: html_notebook
---


```{r}
library(CodeClanData)
```

```{r}
drinks_content
```

#Question 1.

What are the dimensions? 
- [1] 242  18
```{r}
dim(drinks_content)
```
Variable types?
- <chr> characters and <dbl> decimalised numbers
```{r}
str(drinks_content)
```
Variable names, etc.?
- [1] "Beverage_category"       "Beverage"                "Beverage_prep"          
 [4] "Calories"                "Total Fat (g)"           "Trans Fat (g)"          
 [7] "Saturated Fat (g)"       "Sodium (mg)"             "Total Carbohydrates (g)"
[10] "Cholesterol (mg)"        "Dietary Fibre (g)"       "Sugars (g)"             
[13] "Protein (g)"             "Vitamin A (% DV)"        "Vitamin C (% DV)"       
[16] "Calcium (% DV)"          "Iron (% DV)"             "Caffeine (mg)"          
```{r}
names(drinks_content)
```

#Question 2.

Let’s first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.
```{r}
drinks_content %>% 
  select(Beverage_category:Calories)
```

```{r}
drinks_content %>% 
  select(1:4)
```

```{r}
drinks_content %>% 
  select(1:4) %>%
  arrange(Calories, na.rm = TRUE) %>% 
  view()
```

#Question 3.

Create a new variable (column) called calorie_diff, which stores the difference between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. (hint: you’ll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

```{r}
drinks_content %>% 
  mutate(calorie_diff = Calories - 135)
```

#Question 4.

Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

```{r}
  beverages_grouped <- group_by(drinks_content, Beverage_category)
summarise(beverages_grouped, Calories = mean(Calories)) %>% 
  arrange(Calories)
# this took me a while to figure this one out, there has to be better code?
# wanted to use top_n(-3) %>% top_n(3) to only show top and bottom 3 howvever could not achieve?
```

Summary:

As derived from the subset, Coffee, Shaken Iced Beverage and Classic Espresso Drinks have the lowest calorie counts for each category with 4.25, 114.4444 and 140.1724 respectively on average. 
The beverage category's which contain the highest calorie count are Smoothies, Frappuccino® Blended Coffee and Signature Espresso Drinks with 282.2222, 276.9444 and 250 calories respectively.

135 Calories equates to around 10 sugar cubes, therefore in more relatable terms Coffee has around a quarter of a sugar cube, and Smoothies just under 2 full sugar cubes. It is also important to note that there is a large increase in calories from Coffee to the next lowest calorie beverage, more than 110 calories! 

What is not shown however is that Tazo® Tea Drinks have a beverage called Tazo® Tea which specifically has 0 calories within it.

#Question 5.

Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
beverages_grouped <- group_by(drinks_content, Beverage_prep)
summarise(beverages_grouped, Calories = mean(Calories)) %>% 
  arrange(Calories)
  
```

#Question 6.

Which Beverage_prep type contains more than the average calories of all drinks?
Hint
: to answer this, you’ll have to first figure out what the average calories across all drinks are, and then use that as a filter for the grouped Beverage_prep data.

```{r}
drinks_content %>% 
  select(Calories)
```

```{r}
beverages_grouped <- group_by(drinks_content, Beverage_prep)
summarise(beverages_grouped, Calories = mean(Calories)) %>% 
  arrange(Calories)

summarise(drinks_content, Average_Calories_of_all_Beverages = mean(Calories))

filter(beverages_grouped > Average_Calories_of_all_Beverages) %>%
select(Beverage_prep)
 # struggling with this one, dont know how to finish it off and get a summary of Beverage_prep > 193.8719 ?? or to assign the mean function of calories to a variable?
```

The below are Beverage_prep drinks that have a calorie count average which is higher than the average of all drinks:

Soymilk	                207.27273			
Grande Nonfat Milk	    209.61538	
2% Milk	                218.00000			
Venti Nonfat Milk	      260.00000			
Whole Milk	            283.75000	



#Question 7.

Which is the best type of coffee drink to get if you’re worried about consuming too many calories?

```{r}
view(drinks_content)
```

The best Coffee drink when concerned about high calorie intake will be a short Brewed Coffee, which only contains 3 calories.