---
title: "Hypothesis tests & probability "
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)

sleep <- msleep
```

# 1.1 Hypothesis testing - practical
## Question 1.
Explore the dataset and familiarise yourself with it.

```{r}
str(sleep)
head(sleep)

# finding na's
sleep%>% 
  summarise(across(.fns = ~ sum(is.na(.x))))
```

## Question 2.
Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate statistical test to determine whether the mean sleep_total in the sampled population of animal types differs from the typical value for jabberwockies.

*H0: mean sleep of animals = 7*
*Ha: mean sleep of animals ≠ 7*

*Alpha = 0.05*

```{r}
# visualising data
sleep %>% 
  ggplot(
         aes(x = sleep_total)) +
  geom_histogram(col = "white")

sleep %>% 
    ggplot(
    aes(x = sleep_total)) +
  geom_boxplot()
```

```{r}
# observed stat, mean sleep of animals = 10.43 hrs
obs_stat_mean_sleep <- sleep %>% 
  summarise(mean_sleep = mean(sleep_total))
obs_stat_mean_sleep
```

```{r}
# null distribution
null_dist_mean_sleep <- sleep %>% 
  specify(response = sleep_total) %>% 
  hypothesise(null = "point", mu = 7) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
# overlaying observed stat on null dist
# there is a bigger spread here as we have a small sample
null_dist_mean_sleep %>% 
  visualise(bins = 30) +
  shade_pvalue(obs_stat = obs_stat_mean_sleep$mean_sleep,
               direction = "both")

null_dist_mean_sleep %>% 
  summarise(max = max(stat), min = min(stat))
```

```{r}
# calculating and interpreting p-value, p-value < 0
p_value_mean_sleep <- null_dist_mean_sleep %>% 
  get_pvalue(obs_stat = obs_stat_mean_sleep$mean_sleep,
             direction = "both")

p_value_mean_sleep
```

With a p-value less than 0, we reject H0 in favour of Ha, the evidence suggests a significant difference in the null value - the mean sleep of a Jabberwokie can range from 8.845783	to 5.36747, with the mean sleep of the animals within the data set being 10.43373.

## Question 3.
Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

*H0: omnivores_mean_sleep - herbivores_mean_sleep > 0*
*Ha: omnivores_mean_sleep - herbivores_mean_sleep <= 0*

*Alpha = 0.05*

H0: μsleep_total(omni)−μsleep_total(herbi)=0
Ha: μsleep_total(omni)−μsleep_total(herbi)>0
α=0.05

```{r}
# removing na's
sleep_vore <- sleep %>% 
  filter(!is.na(vore)) %>% 
  filter(vore %in% c("herbi", "omni"))

# 9.509375 hrs mean sleep for herbi
mean_herbi_sleep <- sleep_vore %>% 
    filter(vore == "herbi") %>% 
    summarise(mean_herbi_sleep = mean(sleep_total))
mean_herbi_sleep

# 10.925 hrs mean sleep for omni
mean_omni_sleep <- sleep_vore %>% 
  filter(vore == "omni") %>% 
  summarise(mean_omni_sleep = mean(sleep_total))
mean_omni_sleep 
```

```{r}
# visualising data 

# 32 herbi, 20 omni
sleep_vore %>%
  group_by(vore) %>%
  summarise(n = n())

# boxplot
sleep_vore %>%
  ggplot(aes(y = sleep_total, x = vore)) +
  geom_boxplot()

# another style
sleep_vore %>%
  ggplot(aes(x = sleep_total, fill = vore)) +
  geom_boxplot()
```

```{r}
# null dist 
null_dist_vore <- sleep_vore %>% 
  specify(sleep_total ~ vore) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 10000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi")) 
head(null_dist_vore)
```

```{r}
# observed stat= the average difference between omni sleep and herbi sleep is + 1.415625 hrs	
obs_stat_vore <- sleep_vore %>% 
  specify(sleep_total ~ vore) %>%
  calculate(stat = "diff in means", order = c("omni", "herbi"))
  
obs_stat_vore
```

```{r}
# vis the null dist and where the obs stat lies on dist
null_dist_vore %>%
  visualise() +
  shade_p_value(obs_stat = obs_stat_vore, direction = "right")
```

```{r}
# p-value = 0.1239	
p_value_vore <- null_dist_vore %>%
  get_p_value(obs_stat = obs_stat_vore, direction = "right")

p_value_vore
```

With the p-value greater than 0.005, we fail to reject H0 as there is enough data to suggest that omnivores do get on average more sleep than herbivores as there is a statistically significant difference between them.

## Question 4. Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than 5%. 

*H0: prop_dom_animal > 0.05*
*Ha: prop_dom_animal <= 0.05*

*Alpha = 0.05*

H0: πdomesticated=0.05
Ha: πdomesticated>0.05
α=0.05

```{r}
# checking if variance in spelling
sleep %>% 
  distinct(conservation)

# creating categorical variables
sleep_dom <- sleep %>%
  mutate(conservation_flag = ifelse(conservation == "domesticated", 
                               "domesticated", 
                               "not domesticated"))

head(sleep_dom)

```

```{r}
# getting count and proportion of domesticated animals: prop = 12.04819	%
# this 12% is higher than 5% in this small sample, but now checking if statistically higher
sleep_dom %>% 
  group_by(conservation) %>% 
  summarise(count = n(),
            prop = n()/nrow(sleep_dom)) %>% 
  filter(conservation == "domesticated")
```

```{r}
# null dist 
null_dist_dom <- sleep_dom %>%
  specify(response = conservation_flag, success = "domesticated") %>% 
  hypothesize(null = "point", p = 0.05) %>%
  generate(reps = 10000, type = "simulate") %>%
  calculate(stat = "prop") 

head(null_dist_dom)
```

```{r}
# visualising the null dist 
null_dist_dom %>%
  visualise(bins = 20)
```

```{r}
# obs stat = 0.1851852 prop is domesticated animals in our simulation
obs_stat_dom <- sleep_dom %>% 
  specify(response = conservation_flag, success = "domesticated") %>% 
  calculate(stat = "prop")

obs_stat_dom
```

```{r}
# plotting on null dist

null_dist_dom %>%
  visualise(bins = 30) +
  shade_p_value(obs_stat = obs_stat_dom, direction = "right") 
```

```{r}
# p-value = 0.0006
p_value_dom <- null_dist_dom %>%
  get_p_value(obs_stat = obs_stat_dom, direction = "right")

p_value_dom
```


We find the p-value to be less than 0.05, therefore we fail to reject H0, there is enough evidence to suggest that there is more than 5% of domesticated animals in proportion to the other animal types. There is a significant difference, almost around 18.51852% proportion in our simulation.

# 1.2 Hypothesis testing - Defining the Hypothesis

## Question 1.
You work for a independent coffee shop. The boss tells you that she thinks that around 40% of people in the town (population 30,000) know of the coffee shop, but you are skeptical of the estimate. You conduct a random survey of 200 people in the town, asking if respondents are aware of your coffee shop. You want to use the sample to test the hypothesis that 40% or more of the town’s population have heard of the coffee shop.

H0: πawareness=0.4
Ha: πawareness>0.4
*H0: πtown_pop_know_coffeshop >= 0.4*
*Ha: πtown_pop_know_coffeshop < 0.4*

*Alpha = 0.05*

Test
One-sample proportion, right-sided.

Words
H0: the proportion of the town population who are aware of the shop is 40% (or lower)
Ha: the proportion of the town population who are aware of the shop is greater than 40

Null distribution generation
Simulation 


## Question 2.
You work for a website design company and have performed an A/B test on the position of a banner on a website promoting a particular item.

You selected five days at random last month and then randomly selected 200 of each sampled day’s users into group A and another 200 of that day’s users into group B. Group A continued to be shown the banner at the right hand side of the webpage (its usual position) while group B was shown the banner at the top of the page. You monitored each sampled user’s interactions with the website, in particular the ‘click through rate’ (CTR) on the banner, i.e. what proportion of sampled users clicked on the banner. You want to use the sampled data to test the hypothesis that website users overall are more likely to click on the banner if positioned at the top of the page

H0: πCTR(top)−πCTR(right)=0
Ha: πCTR(top)−πCTR(right)>0
*H0: πbanner_click_top - πbanner_click_bottom > 0*
*Ha: πbanner_click_top - πbanner_click_bottom < 0*

*Alpha = 0.05*

Test
Two-samples proportion, right-sided.

Words H0: the CTR with the banner at the top of the website is the same as (or less than) the CTR with the banner at the right hand side of the website.
Ha: the CTR with the banner at the top of the website is greater than the CTR with the banner at the right hand side of the website.

Null distribution generation
Permutation

## Question 3.
You work as an analyst for a car manufacturing company - they have specific standards they must meet for standards and regulation purposes. You have been asked to check the quality control of the manufacture of a particular car part. You have been given data on a sample of 200 parts produced over the period of a week (the sampled parts were pulled at random from the production line and measured: the factory produced a lot more than 200 parts that week). The specification of the part stipulates a width of 145mm, and the manufacturing process is said to have ‘drifted’ if the mean width of parts differs significantly from 145mm. You want to use the sampled measurements to test whether the process overall has drifted.

H0: μwidth=145
Ha: μwidth≠145
*H0: µdiff_in_size = 0*
*Ha: µdiff_in_size ≠ 0*

*Alpha = 0.01*
Test
One-sample mean test, two-sided.

Words
H0: the mean width of the parts produced that week equals 145mm
Ha: the mean width of the parts produced that week differs from 145mm

Null distribution generation
Bootstrap

# 1.3 Hypothesis Testing - Interpreting the results

For the 3 business problems stated above, imagine we performed you got the following p-values (with the given significance levels) write out your interpretation of the results.


## Question 1.

Coffee shop problem. Significance level: 0.05, calculated p-value: 0.07

*P-value is greater than 0.05, therefore we fail to reject H0. This means that there was not enough evidence to suggest a statistically significant difference from the null value. What this means is that from our observations it was more than likely that 40% or more of the towns entire population knew about the coffee shop. Could increase this awareness with a "free morning coffee" event? *


## Question 2.

Website company problem. Significance level: 0.01, p-value: 0.006

*P-value is less than 0.01, therefore we reject H0 in favor of Ha due to sufficient evidence. This means that there was enough evidence in the sample to suggest a statistically significant difference from the null value. This means there was more CTR on the banner on the RHS of the page as opposed to the proposed top positioned banner. Could place desired products/offers within that region on the page for potentially higher awareness?*

## Question 3.

Manufacturing company problem. Significance level: 0.05, -value: 0.55

*P-value more than 0.05, therefore we fail to reject the H0. This means that there was not enough evidence to suggest a statistically significant difference from the null value. This shows that there was no drifiting in the manufacturing process statistically in the volume of parts that were more than the 200 sampled, as the mean width of parts did not differ significantly from 145mm.*