---
title: "Decision trees homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```


<br>
In this homework we will create a decision tree to see which factors are useful in predicting whether or not a passenger on the titanic will survive.  


Run the code below before you begin: 


```{r, warning = FALSE, message = FALSE}
library(rpart)
library(rpart.plot)
library(tidyverse)

titanic_set <- read_csv('data/titanic_decision_tree_data.csv')
shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

<br>

**Data Dictionary**

  * **sex**: Biological Sex, male or female  
  * **age_status**: adult or child (child defined as under 16)  
  * **class** : Ticket class, 1 = 1st (Upper class), 2 = 2nd (Middle Class), 3 = 3rd (Lower Class)    
  * **port_embarkation**: C = Cherbourg, Q = Queenstown, S = Southampton  
  * **sibsp** : number of siblings / spouses aboard the Titanic   
  * **parch**: number of parents / children aboard the Titanic. Some children travelled only with a nanny, therefore parch=0 for them. 
  * **survived_flag** : did they survive, 0 = No, 1 = Yes  



# MVP 


## Question 1  

<br> 
Cleaning up the data is always the first step. Do the following: 

  * Take only observations which have a `survived` flag (i.e. that aren't missing)  
  * Turn your important variables into factors (sex, survived, pclass, embarkation)  
  * Create an `age_status` variable which groups individuals under (and including) 16 years of age into a category called "child" category and those over 16 into a category called "adult".  
  * Drop the NA  
  * Drop any variables you don't need (`X1`, `passenger_id`, `name`, `ticket`, `far`, `cabin`)  

If you need help doing this, the code is below, but please try it yourself first so you can learn!

<br>
<details>
<summary>**Data Cleaning Code** </summary>
<br>

```{r}
titanic_clean <- titanic_set %>%
  filter(survived %in% c(0,1)) %>%
# Convert to factor level
	mutate(sex = as.factor(sex), 
	       age_status = as.factor(if_else(age <= 16, "child", "adult")),
         class = factor(pclass, levels = c(3,2,1), labels = c("Lower", "Middle", "Upper")), 
	       survived_flag = factor(survived, levels = c(0,1), labels = c("No", "Yes")), 
	       port_embarkation = as.factor(embarked)) %>%
  select(sex, age_status, class, port_embarkation, sib_sp, parch, survived_flag) %>%
  na.omit()
```
</details>


<br>

## Question 2  

<br> 
Have a look at your data and create some plots to ensure you know what you're working with before you begin. Write a summary of what you have found in your plots. Which variables do you think might be useful to predict whether or not people are going to die? Knowing this before you start is the best way to have a sanity check that your model is doing a good job.  


```{r}
titanic_clean %>%
  ggplot(aes(x = survived_flag)) +
  geom_bar()
```

```{r}
titanic_clean

# check the effect of sex 
titanic_clean %>%
  select(survived_flag, sex) %>%
  group_by(survived_flag,sex) %>%
  summarise(total_count = n()) %>%
  ggplot() + 
  aes(x = survived_flag, y = total_count, fill = sex) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```


```{r}
titanic_clean %>%
  select(survived_flag, class) %>%
  group_by(survived_flag,class) %>%
  summarise(total_count = n()) %>%
  ggplot() + 
  aes(x = survived_flag, y = total_count, fill = class) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```

```{r}
titanic_clean %>%
  select(survived_flag, age_status) %>%
  group_by(survived_flag,age_status) %>%
  summarise(total_count = n()) %>%
  ggplot() + 
  aes(x = survived_flag, y = total_count, fill = age_status) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```


```{r}
titanic_clean %>%
  select(survived_flag, port_embarkation) %>%
  group_by(survived_flag,port_embarkation) %>%
  summarise(total_count = n()) %>%
ggplot() + 
  aes(x = survived_flag, y = total_count, fill = port_embarkation) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```

My explorations tell me that probably sex, class, age, and port embarkation are going to be useful in predicting whether or not someone will survive because there seems to be differences between these.   



<br>

## Question 3  

<br> 
Now you can start to build your model. Create your testing and training set using an appropriate split. Check you have balanced sets. Write down why you chose the split you did and produce output tables to show whether or not it is balanced. [**Extra** - if you want to force balanced testing and training sets, have a look at the `stratified()` function in package `splitstackshape` (you can specify multiple variables to stratify on by passing a vector of variable names to the `group` argument, and get back testing and training sets with argument `bothSets = TRUE`)]

```{r}
# set the random seed number: ONLY INSTRUCTORS - students SHOULD NOT do this in their homework 
set.seed(3)

n_data <- nrow(titanic_clean)

# create a test sample index
test_index <- sample(1:n_data, size = n_data*0.2)

# create test set
titanic_test  <- slice(titanic_clean, test_index)

# create training set
titanic_train <- slice(titanic_clean, -test_index)

# check balanced sets
titanic_test %>%
 janitor::tabyl(survived_flag)

titanic_train %>%
  janitor::tabyl(survived_flag)
```



If we like, we can **force stratification** on the `survived_flag` variable using the `stratified()` function from package `splitstackshape`. In fact, let's go further and force stratification on `survived_flag`, `sex`, `class` and `age_status`. The more you stratify, however, the more difficult it will be to enforce balance between the testing and training sets for small datasets, as you start to 'run out' of rows for each distinct groups.

```{r}
library(splitstackshape)

sets <- stratified(titanic_clean, group = c("survived_flag", "sex", "class", "age_status"), bothSets = TRUE, size = 0.2)

titanic_test <- sets$SAMP1
titanic_train <- sets$SAMP2

nrow(titanic_test)
nrow(titanic_train)

titanic_test %>%
 janitor::tabyl(survived_flag)
titanic_train %>%
  janitor::tabyl(survived_flag)
```

```{r}
titanic_test %>%
  janitor::tabyl(survived_flag, sex) %>%
  janitor::adorn_percentages()

titanic_train %>%
  janitor::tabyl(survived_flag, sex) %>%
  janitor::adorn_percentages()
```

```{r}
titanic_test %>%
  janitor::tabyl(survived_flag, class) %>%
  janitor::adorn_percentages()

titanic_train %>%
  janitor::tabyl(survived_flag, class) %>%
  janitor::adorn_percentages()
```

```{r}
titanic_test %>%
  janitor::tabyl(survived_flag, age_status) %>%
  janitor::adorn_percentages()
titanic_train %>%
  janitor::tabyl(survived_flag, age_status) %>%
  janitor::adorn_percentages()
```


## Question 4      

<br> 
Create your decision tree to try and predict survival probability using an appropriate method, and create a decision tree plot.

```{r}
titanic_fit <- rpart(survived_flag ~ ., 
                     data = titanic_train, 
                     method = 'class')

rpart.plot(titanic_fit, yesno = 2)
```
<br>

```{r}
rpart.plot(titanic_fit, yesno = 2, type = 4, extra = 101)
```

## Question 5    

<br> 
Write down what this tells you, in detail. What variables are important? What does each node tell you? Who has the highest chance of surviving? Who has the lowest? Provide as much detail as you can.    

The trees that you calculate will all differ, but a good way to approach this would be to work in descending order of group size, stating the conditions fulfilled by that group and the conditions required to reach that group. For example:

* The largest group in the data (43%) comprises adult, male, lower and middle class passengers with a probability of survival of 0.11
* The next largest group (22%) comprises female middle and upper class passengers, with a probability of survival of 0.94
* and so on

<br>

## Question 6     

<br>  
Test and add your predictions to your data. Create a confusion matrix. Write down in detail what this tells you for this specific dataset.  

```{r}
library(modelr)
library(yardstick)
library(caret)

# add the predictions
titanic_test_pred <- titanic_test %>%
                 add_predictions(titanic_fit, type = 'class')


conf_mat <- titanic_test_pred %>%
              conf_mat(truth = survived_flag, estimate = pred)
conf_mat
```

Let's also see accuracy, TPR (sensitivity) and TNR (specificity)

```{r}
titanic_test_pred %>%
  accuracy(truth = survived_flag, estimate = pred)
```
```{r}
titanic_test_pred %>%
  yardstick::sensitivity(truth = survived_flag, estimate = pred)
```

```{r}
titanic_test_pred %>%
  yardstick::specificity(truth = survived_flag, estimate = pred)
```

Or using `caret`:

```{r}
library(caret)
confusionMatrix(data = titanic_test_pred$pred, reference = titanic_test_pred$survived_flag)  
```


# Extension  

See how a `ranger()` random forest classifier compares with a single decision tree in terms of performance. Can you tune the values of the `mtry`, `splitrule` and `min.node.size` hyperparameters? Which variables in the dataset turn out to be most important for your best model? The `Kappa` metric might be the best one to focus on if you want to improve performance for an imbalanced data set. Do some research on the definition of `Kappa` before you start.

We provide the code in the dropdown below if you get stuck, but still want to play around with this (note that run time can be up to 5-10 mins for the tuning). **Save your notebook before you begin** in case you need to force quit your session!

<br>
<details>
<summary>**Code**</summary>

```{r}
library(ranger)

control <- trainControl(
  method = "repeatedcv", 
  number = 5, 
  repeats = 10
)

tune_grid = expand.grid(
  mtry = 1:6,
  splitrule = c("gini", "extratrees"),
  min.node.size = c(1, 3, 5)
)
```

```{r}
rf_tune <- train(
  survived_flag ~ ., 
  data = titanic_train, 
  method = "ranger",
  metric = "Kappa",
  num.trees = 1000,
  importance = "impurity",
  tuneGrid = tune_grid, 
  trControl = control
)

plot(rf_tune)
rf_tune
```
</details>
<br>

Here is the highest performing tuned model obtained from one of the tuning runs (note, this is also a random process, so the hyperparameters we choose below may not be the ones suggested by the data above!)

```{r}
rf_classifier <- ranger(survived_flag ~ ., mtry = 3, data = titanic_train, splitrule = "gini", min.node.size = 5, num.trees = 1000, importance = "impurity")
```

How do we do on the training set?

```{r}
titanic_train_pred <- titanic_train %>%
  mutate(pred = predict(rf_classifier, data = titanic_train)$predictions)

confusionMatrix(data = titanic_train_pred$pred, reference = titanic_train_pred$survived_flag)
```

Next, how do we do on the test set?

```{r}
titanic_test_pred <- titanic_test %>%
  mutate(pred = predict(rf_classifier, data = titanic_test)$predictions)

confusionMatrix(data = titanic_test_pred$pred, reference = titanic_test_pred$survived_flag)
```

Finally, which variables are most important?

```{r}
importance(rf_classifier)
```

