---
title: "my_ans_ggplot_homework_2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(CodeClanData)
library(tidyverse)

backpack <- backpack
```

### Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.
```{r}
view(backpack)
glimpse(backpack)
```

```{r}
theme<- 
  theme(
    text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dashed")
  )
```

### Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

A. The three most common backpack weights are 5, 8 and 10 lbs with the average being around 12 lbs. 
```{r}
ggplot(backpack,
  aes(x = backpack_weight)) +
  geom_bar() +
  ggtitle("\nDistribution of backpack weight carried by students\n") +
  xlab("\nWeight of backpack, lbs") +
  ylab("Number of Students\n") +
  scale_y_continuous(breaks = 0:12) +
  geom_vline(xintercept = mean(backpack$backpack_weight), 
             color="red") +
  theme(plot.title = element_text(hjust = 0.5)) +
theme
```

### Question 3.

Are there differences in backpack weight between sex?

A. Females mostly carry 10lbs backpacks and males 11lbs,  however more females can be seen as carrying on average a heavier weight than males.
```{r}
ggplot(backpack,
  aes(x = backpack_weight, fill = sex)) +
  geom_bar() +
  ggtitle("\nDistribution of backpack weight carried by students\n") +
  xlab("\nWeight of backpack, lbs") +
  ylab("Number of Students\n") +
  labs(fill = "Sex") +
  theme(plot.title = element_text(hjust = 0.5)) +
theme
```


### Question 4.

What is the relationship between body weight and backpack weight?

A. The heavier the student gets, the heavier the backpack becomes up until around the 200lbs mark, where backpack weight actually starts to decrease. This could be due to an increase in strength with overall size. However at the 200lbs+ mark for body weight there could be an element of obesity and therefore lack of general fitness and strength - this would potentially explain the change in trend.
```{r}
ggplot(backpack,
       aes(x = body_weight, y = backpack_weight)) +
  geom_point(colour = "blue") +
  geom_smooth() +
  ggtitle("\nBodyweight vs Backpack weight\n") +
  xlab("\nBodyweight, lbs") +
  ylab("Weight of backpack, lbs\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme
```

### Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

A. The relationship appears to be proportionally similar.
```{r}
ggplot(backpack,
       aes(x = body_weight, y = backpack_weight, fill = sex)) +
  geom_col() +
  ggtitle("\nBodyweight vs Backpack weight",
          subtitle = "Using sex as a factor") +
  xlab("\nBodyweight, lbs") +
  ylab("Weight of backpack, lbs\n") +
  labs(fill = "Sex") +
  theme
```


### Question 6.

What proportion of students report back pain?

A. 68% of students do not report back pain compared to the 32% that do.

```{r}
backpack %>% 
  mutate(back_problems = if_else(back_problems == 1,
                                 "Yes",
                                 "No")) %>% 
ggplot(aes(x = back_problems, fill = back_problems)) +
  geom_bar()+
  ggtitle("\nBack problems in students",
          subtitle = "Proportion out of 100 students") +
  xlab("") +
  ylab("Number of students\n") +
  labs(fill = "Back problems")+
  scale_fill_manual(values = c('grey40', 'red'),
                    guide = "none") +
  geom_text(aes(label = ??????),
            colour = "black", size = 2, hjust = 1) +
  theme
```

### Question 7.

Is this proportion different between men and women?

A. 36/68 that reported no back problems are male, which is 52.94% - this shows no experience of back problems between genders is around the same. 
   8/32 that reported back problems are male, equating to 25%. Out of those experiencing back problems there is a significantly higher number of females in    proportion, in fact there a re 3 times more females with back pain than males.
```{r}
backpack %>% 
  mutate(back_problems = if_else(back_problems == 1,
                                 "Yes",
                                 "No")) %>% 
ggplot(aes(x = back_problems, fill = sex)) +
  geom_bar()+
  ggtitle("\nBack problems in students",
          subtitle = "Proportion out of 100 students using sex as a factor") +
  xlab("") +
  ylab("Number of students\n") +
  labs(fill = "Back problems") +
  scale_fill_manual(values = c('red', 'blue')) +
  theme

36*100/68
8*100/32
```


### Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported back pain.

Try both backpack_weight and body_weight as well as the ratio.

```{r}
backpack %>% 
  mutate(back_problems = if_else(back_problems == 1,
                                 "Yes",
                                 "No")) %>% 
ggplot(aes(x = body_weight, y = backpack_weight, fill = sex)) +
  geom_col() +
  facet_grid(~back_problems) +
  ggtitle("\nBodyweight vs Backpack weight",
          subtitle = "Using sex and back problems as factors") +
  xlab("\nBodyweight, lbs") +
  ylab("Weight of backpack, lbs\n") +
  labs(fill = "Sex") +
  scale_fill_manual(values = c('red', 'blue')) +
  theme
```

