---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 
```{r}
library(tidyverse)
library(janitor)
library(broom)
library(factoextra)
```

```{r}
customers <- read_csv("data/mall_customers.csv") %>% 
  clean_names()
```

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 
```{r}
customers_money <- customers %>%
  select(spending_score_1_100, annual_income_k)

customers_money %>% 
  ggplot() +
  aes(x = annual_income_k,
      y = spending_score_1_100) +
  geom_point()

customers_money <- customers_money %>%
              mutate_if(is.numeric, scale)

fviz_nbclust(customers_money,
             kmeans,
             method = "wss",
             nstart = 25)

fviz_nbclust(customers_money,
             kmeans,
             method = "silhouette",
             nstart = 25)

fviz_nbclust(customers_money,
             kmeans,
             method = "gap_stat",
             nstart = 25,
             k.max = 10)
```

* Perform k-means clustering and chose a value of k.

```{r}
max_k <- 25

k_cluster_customer <- tibble(k = 1:max_k) %>% 
  mutate(
    kclust = map(k, ~ kmeans(customers_money, .x, nstart = 25)),
    tidied = map(kclust, tidy), 
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers_money)
    )

clustering <- k_cluster_customer %>% 
  unnest(glanced)

clustered_customer <- kmeans(customers_money,
                        centers = 5, # how many clusters
                        nstart = 25)


```


* Visualise the clustering for your chosen value of k.

```{r}
clustering %>% 
  unnest(augmented) %>% 
  filter(k == 5) %>% 
  ggplot() +
  aes(x = spending_score_1_100,
      y = annual_income_k,
      colour = .cluster) +
  geom_point()
```


* Do you think the clustering seems a good fit for this data?

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 






