---
title: "Tidyverse consolidation homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(tidyr)
library(stringr)
```


# MVP 

<br>

---**Question 1.**  
Load the `code_clan_tweets.csv` data. Find the number of rows, columns, and list all the variable names. 

```{r}
code_clan_tweets <- janitor::clean_names(read_csv("data/code_clan_tweets.csv")) 

code_clan_tweets %>% 
  dim()

code_clan_tweets %>% 
  names()

```


<br>

---**Question 2.**  
Find the total number of favourites (stored in `favorite_count`) that CodeClan tweets have got. Don't add a tweet's `favorite_count` to the total when the tweet was quoted (see the `is_quote` variable). 

```{r}
code_clan_tweets %>% 
  filter(is_quote == FALSE) %>% 
  summarise(favorite_count = sum(favorite_count))
```


<br>

---**Question 3.**  
Summarise the mean number of retweets (`retweet_count`), split by the type of platform that was used to tweet (`source`). Again, omit quoted tweets from the mean.

```{r}
code_clan_tweets %>% 
  filter(is_quote == FALSE) %>%
  group_by(source) %>% 
  summarise(retweet_count = mean(retweet_count))
```


<br>

---**Question 4.**  
Count the total number of likes (i.e. total of `favorite_count`), split by `media` type, and arrange them from most likes to least. Assume that any tweets without a listed `media` type are of type "text".  

```{r}
code_clan_tweets %>%
  group_by(media_type) %>% 
  summarise(num_of_likes = sum(favorite_count))
```

```{r}
#instructor answer
code_clan_tweets %>%
  mutate(media_type = coalesce(media_type, "text")) %>%
  group_by(media_type) %>%
  summarise(favourite_count_total = sum(favorite_count)) %>%
  arrange(desc(favourite_count_total))
```


<br>

---**Question 5.**  
Find the mean number of characters that a CodeClan tweet contains.  You can either find a way to count the text in the `text` variable, or use the variable `display_text_width` which contains a count. Checking both would be a good idea if possible to see if they match (but it's not essential).  
<br>

```{r}
code_clan_tweets %>%
  summarise(avg_tweet_length = mean(display_text_width))
```

```{r}
code_clan_tweets %>%
  summarise(avg_tweet_length = mean(display_text_width))
```
```{r, warning = FALSE}
# alternative way of testing the str_length function
code_clan_tweets_text <- code_clan_tweets %>%
  mutate(count_str_length = str_length(text)) %>%
  select(count_str_length, display_text_width)

code_clan_tweets_text
```

---**Question 6.**  
The `code_clan_info.csv` data file contains `status_url` along with other info. Load this in, and join it to the `code_clan_tweets` tibble, so that you have a `status_url` for each tweet. Decide which variable to join the tibbles on.  

```{r}
code_clan_info <- janitor::clean_names(read_csv("data/code_clan_info.csv"))
```

```{r}
code_clan_info
code_clan_tweets
cc_tweet_data <- full_join(code_clan_info, code_clan_tweets, by = "tweet_id")
```

<br>

---**Question 7.**  
From your new joined data, create a new tibble `codeclan_hashtags` containing only the `tweet_id` and convert the hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- cc_tweet_data %>%
  filter(hashtags != is.na(hashtags)) %>%
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  group_by(tweet_id) %>%
  summarise(hashtags)
codeclan_hashtags
```

```{r}
# instructors answers: create a variable which has all the hashtags in lowercase for analysis
codeclan_hashtags_inst <- codeclan_hashtags%>% 
  select(tweet_id, hashtags) %>%
  mutate(lowcase_hashtag = str_to_lower(hashtags)) %>%
  select(-hashtags) %>%
  drop_na(lowcase_hashtag)

codeclan_hashtags_inst
```

<br>

# Extensions

<br>

---**Question 8.**
Some tweets have more than one hashtag, denoted by the `c(` symbols at the start of the string. Use the `str_sub()` and `str_detect()` functions and your `codeclan_hashtags` tibble from above to find all the cases in which the hashtag string begins with characters`c(`.
<details>
<summary>**Hint**</summary>
Try a regex pattern `"c\\("`. The double backslash is necessary in R to mark a backslash as itself being escaped. 
</details>

<details>
<summary>**Answer**</summary>
```{r}
# 8 : use the str_sub function to get the first 2 letters of each tweet
hashtags_multiple <- codeclan_hashtags %>% 
  mutate(first_letters = str_sub(lowcase_hashtag, 1, 2)) %>%
  filter(str_detect(first_letters, "c\\("))

hashtags_multiple
```
</details>

<br>

---**Question 9.**
Use the `str_detect()` function to find all tweets with `text` that mentions "Edinburgh", and count how many such tweets there are. 
<details>
<summary>**Hint**</summary>
You may need to make sure your tweets are all the same case before searching.
</details>

<details>
<summary>**Answer**</summary>
```{r}
# 9: Use the str_detect() function to find all cases where a tweet text mentions edinburgh, and count in how many tweets it appears
codeclan_tweets %>%
  mutate(lowcase_tweets = str_to_lower(text)) %>%
  filter(str_detect(lowcase_tweets, "edinburgh")) %>%
  summarise(count = n())
```
</details>

<br>

---**Question 10.**
Use `str_extract_all()` with a regular expression to find out which Twitter users CodeClan have been tweeting. 
<details>
<summary>**Hint**</summary>
Think about all the different characters which could be in a Twitter username, do a bit of online research
</details>

<details>
<summary>**Answer**</summary>
```{r}
# 10: use str_extract_all and regex to find usernames
user_pattern <- "@[a-zA-Z0-9_]+"
tweets <- codeclan_tweets %>% 
  select(text)
head(str_extract_all(tweets$text, user_pattern))
```
</details>