---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>
```{r}
library(tidyverse)
library(fastDummies)
library(mosaic)
library(mosaicData)
library(GGally)
library(ggfortify)
library(janitor)
```

# MVP

1. Load the `diamonds.csv` data set and undertake an initial exploration of the data. You will find a description of the meanings of the variables on the relevant [Kaggle page](https://www.kaggle.com/shivam2503/diamonds/)

```{r}
diamond <- read_csv("data/diamonds.csv") %>% 
  clean_names()

# no na's
summary(diamond)
```

2. We expect the `carat` of the diamonds to be strong correlated with the physical dimensions `x`, `y` and `z`. Use `ggpairs()` to investigate correlations between these four variables.

```{r}
diamonds_samp <- diamond[sample(1:dim(diamonds)[1], 1000), ]

ggpairs(diamonds_samp[, c(2, 9:11)],
 title = "Diamonds"
)
```

3. So, we do find significant correlations. Let's drop columns `x`, `y` and `z` from the dataset, in preparation to use only `carat` going forward.
```{r}
diamond_trim <- diamond %>% 
  select(-c(x, y, z))
diamond
```

4. We are interested in developing a regression model for the `price` of a diamond in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `price` and the predictors (this may take a while to run, don't worry, make coffee or something).

  ii. Perform further `ggplot` visualisations of any significant correlations you find.
  
```{r}
ggpairs(diamond_trim)
```

```{r}
diamond_trim %>%
  ggplot(aes(x = price, y = carat)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

diamond_trim %>%
  ggplot(aes(x = table, y = depth)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

5. Shortly we may try a regression fit using one or more of the categorical predictors `cut`, `clarity` and `color`, so let's investigate these predictors. Investigate the levels of these predictors. How many dummy variables do you expect for each of them?

```{r}
alias(lm(price ~ ., data = diamond_trim))


```


6. Start with simple linear regression. Regress `price` on `carat` and check the regression diagnostics.

```{r}
model <- lm(price ~ carat, data = diamond_trim)

summary(model)

autoplot(model)

diamond_trim %>%
  ggplot(aes(x = price, y = carat)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
model2 <- lm(price ~ carat + color,
             data = diamond_trim)

summary(model2)

autoplot(model2)
```


# Extension
    
8. Try adding an interaction between `log(carat)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction
    