---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(CodeClanData)
```

```{r}
revenue <- clean_names(qb_revenue_breakdown)
monthly_sales <- clean_names(qb_monthly_sales)
competitors <- clean_names(qb_competitors)
device_data <- clean_names(qb_device_data)
```

# 2.1 Part 1

### Question 1.
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(revenue,
       aes(x = product, y = revenue,
           fill = year)) +
  geom_col()
```

### Question 1.2
```{r}
ggplot(revenue,
       aes(x = product, y = revenue,
           fill = year)) +
  geom_col() +
  theme_minimal() +
  labs(
    title = "Revenue Streams",
    subtitle = "2018-2019",
    x = "\nProduct",
    y = "Revenue, Currency\n",
    fill = "Year"
  )+
  theme(axis.text = element_text(size = 12))+
  scale_fill_manual(
    values = col_scheme
    )
```

### Question 2.
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}

# there is a quicker way to do this

monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
ggplot(aes(x = date,
           y = amount)) +
  geom_line() +
  theme_minimal()
```

### Question 2.1
```{r}
monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
ggplot(aes(x = date,
           y = amount)) +
  geom_line() +
  scale_fill_manual(
    values = col_scheme
    )+
  theme_minimal()+
   theme(axis.text = element_text(size = 12))
  
  
```


### Question 3.
Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
monthly_sales %>% 
  filter(cashflow != "Kwikbit Unit 1 Sales") %>% 
ggplot(aes(x = date,
           y = amount,
           fill = cashflow)) +
  geom_line()+
  geom_point(aes(fill = cashflow))
```

### Question 3.2
```{r}
monthly_sales %>% 
  filter(cashflow != "Kwikbit Unit 1 Sales") %>% 
ggplot(aes(x = date,
           y = amount,
           fill = cashflow)) +
  geom_line()+
   scale_fill_manual(
        values = col_scheme) +
  geom_point()+
  theme_minimal()
```


### Question 4.
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
ggplot(competitors,
       aes(x = year,
           y = revenue,
           fill = company)) +
  geom_area()
```

### Question 4.2
```{r}
ggplot(competitors,
       aes(x = year,
           y = revenue,
           fill = company)) +
  geom_area()+
   scale_fill_manual(
        values = col_scheme) +
  labs(
    title = "Annual Sales over 5 Years",
    x = "",
    y = "Revenue, Currency",
    fill = "Company"
  )+
  scale_y_continuous(labels = scales::comma) +
  theme_minimal()
```

### Question 5.
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(competitors,
       aes(x = year,
           y = revenue,
           colour = company))+
  geom_line()+
  geom_point()
```

### Question 5.2
```{r}
ggplot(competitors,
       aes(x = year,
           y = revenue,
           colour = company))+
  geom_line()+
  labs(
    title = "Annual Sales over 5 Years",
    x = "",
    y = "Revenue, Currency",
    fill = "Company")+
  geom_point() +
       scale_fill_manual(
        values = col_scheme) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal()
```


### Question 6.
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data. 
```{r}
device_data %>% 
  mutate(mins = hours*60 + mins) %>% 
ggplot(aes(x = mins,
           y = counts))+
  geom_line()+
  theme(axis.text = element_text(size = 12))+
  scale_fill_manual(values = col_scheme)+
  facet_wrap(~device)+
  theme_minimal()
```

## 2.2 Part 2

### Question 1.
Take the plots that you produced in part one and now polish them by:

  Adding appropriate labels
  Changing the scales and coordinates when appropriate.
  Applying a unified theme, which is described below:
  Graphs should have white backgrounds, and use colour sparingly.
  There should be faint grid lines.
  Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
    All plots should use colours from the following company colour scheme.

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```

### Question 1.2
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(revenue,
       aes(x = product, y = revenue,
           fill = year)) +
  geom_col() +
  theme_minimal() +
  labs(
    title = "Revenue Streams",
    subtitle = "2018-2019",
    x = "\nProduct",
    y = "Revenue, Currency\n",
    fill = "Year"
  )+
  theme(axis.text = element_text(size = 12))+
  scale_fill_manual(values = col_scheme
  )
```