---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
library(janitor)
```

```{r}
titanic <- read_csv("data/titanic.csv")
```

## 1.1 Titanic

```{r}
# creating crew column identifying status of passengers/crew
titanic <- titanic %>%
  mutate(
    crew = case_when(
      str_detect(class, "crew")  ~ TRUE,
      str_detect(class, "staff") ~ TRUE,
      TRUE ~ FALSE
    )
  )
```

```{r}
# getting count for crew and passengers
titanic_crew <- titanic %>% 
  filter(crew == TRUE)

titanic_passenger <- titanic %>% 
  filter(crew == FALSE)
```

A: 0.2370787 / 23.71% / 1 in 4.218009 survived
```{r}
# dividing number of crew survived by total number of crew
prob_crew_survive <- titanic_crew %>% 
  filter(crew == TRUE, survived == "yes") %>% 
  summarise(prob_crew_survive = n() / nrow(titanic_crew))

prob_crew_survive

# number of survivors
0.2370787 * 890

# simplifying fraction
890/211
```

A: 0.3796507	/ 37.97% / 1 in 2.634 passengers survived.
```{r}
# dividing number of passengers survived by total number of passengers
prob_passenger_survive <- titanic_passenger %>% 
  filter(crew == FALSE, survived == "yes") %>% 
  summarise(prob_passenger_survive = n() / nrow(titanic_passenger))

prob_passenger_survive

# number of survivors
0.3796507	 * 1317

# simplifying fraction
1317/500
```

## 1.2 Probability Consolidation Questions

### Question 1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

- A: 0.6666667 / 66% / 2 in 3 rolls
```{r}
# rolling one dice and exploring potential outcomes
one_dice <- rolldie(1) 

# filtering above results to specifically less than 5
dice_less_five <- one_dice %>% 
  filter(X1 < 5)

# dividing number of results less than 5 with number of total unfiltered results
prob_less_five <- nrow(dice_less_five) / nrow(one_dice)
prob_less_five
```

### Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?

A: 0.01851852 / 1.8% / 19 in 64 rolls
```{r}
# rolling three die and exploring potential outcomes
three_dice <- rolldie(3) 

# filtering above results to specifically less than 5
three_die_less_five <- three_dice %>% 
  filter(X1 < 5, X2 <5, X3 <5)

# dividing number of results less than 5 with number of total unfiltered results
three_die_prob_less_five <- nrow(three_die_less_five) / nrow(three_dice)
three_die_prob_less_five

# gives how many rolls out of 64 to achieve less than 5
0.2962963 * 64

```

### Question 3.
In an experiment where we roll four dice:

- how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?
A: 34 Outcomes

- what is the probability of this event?
A: 0.02623457 / 2.6% / 34 in 1296 rolls / 1 in 38.11765 rolls

- what assumptions do we apply in answering part ii?
A: ????


```{r}
# rolling three die and exploring potential outcomes
four_die <- rolldie(4) 

# filtering above results and creating new col with count of die rolls equal to desired outcomes
four_die_outcome <- four_die %>% 
  mutate(total = rowSums(.)) %>% 
  filter(total %in% c(5,6,7))

# dividing number of results of intended outcomes with number of total unfiltered results
prob_four_die_outcome <- nrow(four_die_outcome) / nrow(four_die)
prob_four_die_outcome

# gives how many rolls out of 1296 to achieve intended outcomes
0.02623457* nrow(four_die)

# simplifying fraction
nrow(four_die)/34

```

### Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

A: 0.02 / 2% / 1 in 50 attempts
```{r}
# n(total) = n1 x n2 x n3.... x nn
outcomes <- 5  * 5 * 2
outcomes

prob <- 1/outcomes
prob
```

