---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
library(gtools)
library(janitor)
library(infer)
```

## Task 1.
Load the data again, clean_names(), and re-familiarise yourself with it
```{r}
ames <- clean_names(read_csv("data/ames.csv"))
glimpse(ames)
```


## Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
# disttribution is not normal, it is extremely skewed right as there are some but very few high lot_area scores.
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(col = "white",
                 fill = "#66ff66",
                 alpha = 0.4)
```


## Task 3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.
```{r}
boostrap_sample_200 <- ames %>% 
  rep_sample_n(size = 200, replace = TRUE, reps = 5000) %>% 
  summarise(mean_lot_area = mean(lot_area))

boostrap_sample_200 %>% 
  summarise(mean(mean_lot_area))
```

```{r}
ggplot(boostrap_sample_200) +
  aes(x = mean_lot_area) +
geom_histogram(col = "white",
               fill = "#66ff66",
               alpha = 0.4)

boostrap_sample_200 %>% 
  summarise(sd_bs_mean_lot_area = sd(mean_lot_area))
```

## Task 4.
Use your bootstrap distribution to calculate a 95%

CI for mean(lot_area), and visualise it on the distribution

```{r}
infer_resample <- boostrap_sample_200 %>% 
  specify(response = mean_lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
infer_ci_95 <- infer_resample %>% 
  get_confidence_interval(level = 0.95,
                          type = "percentile")
infer_ci_95

infer_ci_95 <- infer_resample %>% 
  get_ci(level = 0.95,
                          type = "percentile")
infer_ci_95
```

```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci_95)
```

## Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99%
CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95%
CI? Does that make sense?
```{r}
# it is broader, as would be expected as there is now a wider "net" of distribution
ci_99 <- boostrap_sample_200 %>% 
  summarise(mean = mean(mean_lot_area),
            lower_bound = quantile(mean_lot_area, probs = 0.005),
            upper_bound = quantile(mean_lot_area, probs = 0.995))

ci_95 
```


## Task 6.
Calculate the point estimate of the mean(lot_area)